MCC Models

### 2019-03-06 papers/full_paper/models.Rmd
adult models for paper first pass

---
title: "MCC Model Predictions"
author: "Manuel Bohn"
date: "15 6 2018"
output: html_document
---

@@

## Adult model predictions

@@

```{r model setup, include = FALSE}

@@

# combining all priors into one data frame
predPrior = rbind(strongPrior,mediumPrior,weakPrior)

# data frame to keep track of unique conditions
allConditions <- predPrior %>%
  select(-prior)

@@

#### Pragmatic model predictions

```{r model rsa noise, include = FALSE}
predictions.rsa <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

@@

# ex4_pragm_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditions, predPrior = predPrior), 
#   data_var = "dataFromR",
#   model_var = "model",
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model.rds")

ex4_pragm_model <- readRDS("../../stats/saves/ex4_pragm_model.rds")

# model predictions
ex4_pragm_model_pred <- ex4_pragm_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(model="pragmatic",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))

## noise parameter
ex4_pragm_model_noise <- ex4_pragm_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

#### Prior only model predictions

```{r model prior only , include = FALSE}
predictions.prior.only <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

@@

```{r predictions prior only model, include = FALSE}

# ex4_prior_only_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditions, predPrior = predPrior),
#   data_var = "dataFromR",
#   model_var = "model",
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model.rds")

ex4_prior_only_model <- readRDS("../../stats/saves/ex4_prior_only_model.rds")

# model predictions
ex4_prior_only_model_pred <- ex4_prior_only_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(model="prior_only",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))

## noise parameter
ex4_prior_only_model_noise <- ex4_prior_only_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

```{r predictions prior only model, include = FALSE}

>>>>

---
title: "MCC Models"
author: "Manuel Bohn"
output: html_document
---

@@

## Adult models

### Experiment 3
#### Model parameters and data

```{r ex3 data}
# load data  from experiment 4 to infer noise parameter
adult_ex3_data <-read_csv(file="../../stats/data/adult_ex3.csv") 
```

```{r}
# load data from prior manipulation experiments to set priors
adult_ex2_novelty_data <- read_csv(file="../../stats/data/adult_ex2_novelty.csv")
adult_ex2_preference_data <- read_csv(file="../../stats/data/adult_ex2_preference.csv")
```

```{r model setup ex 4, include = FALSE}

# create datafiles with model parameters
predPriorEx3 <- data.frame(
  common_ground_manipulation = c("preference","preference","preference","preference","novelty","novelty","novelty","novelty"),
  speaker = c("same_speaker","same_speaker","different_speaker","different_speaker","same_speaker","same_speaker","different_speaker","different_speaker"),
  alignment = c("congruent","incongruent","congruent","incongruent","congruent","incongruent","congruent","incongruent"))

prior_pref_same <- adult_ex2_preference_data %>%
  filter(condition == "same_speaker")%>%
  summarise(prior = mean(correct))

prior_pref_diff <- adult_ex2_preference_data %>%
  filter(condition == "different_speaker")%>%
  summarise(prior = mean(correct))

prior_novel_same <- adult_ex2_novelty_data %>%
  filter(condition == "same_speaker")%>%
  summarise(prior = mean(correct))

prior_novel_diff <- adult_ex2_novelty_data %>%
  filter(condition == "different_speaker")%>%
  summarise(prior = mean(correct))

predPriorEx3$prior = list(
    c(1-prior_pref_same$prior,1-prior_pref_same$prior,prior_pref_same$prior), 
    c(prior_pref_same$prior,prior_pref_same$prior,1-prior_pref_same$prior), 
    c(1-prior_pref_diff$prior,1-prior_pref_diff$prior,prior_pref_diff$prior), 
    c(prior_pref_diff$prior,prior_pref_diff$prior,1-prior_pref_diff$prior),
    c(1-prior_novel_same$prior,1-prior_novel_same$prior,prior_novel_same$prior), 
    c(prior_novel_same$prior,prior_novel_same$prior,1-prior_novel_same$prior), 
    c(1-prior_novel_diff$prior,1-prior_novel_diff$prior,prior_novel_diff$prior), 
    c(prior_novel_diff$prior,prior_novel_diff$prior,1-prior_novel_diff$prior))

# data frame to keep track of unique conditions
allConditionsEx3 <- predPriorEx3 %>%
  select(-prior)

#### Model predictions

##### Pragmatic model

```{r model rsa noise ex3, include = FALSE}
predictions.rsa.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

```{r predictions pragmatic model ex 3, include = FALSE}

# ex3_pragm_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model.rds")

ex3_pragm_model <- readRDS("../../stats/saves/ex3_pragm_model.rds")

# model predictions
ex3_pragm_model_pred <- ex3_pragm_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation","speaker", "alignment"), sep="/")%>%
  mutate(model="pragmatic")

## noise parameter
ex3_pragm_model_noise <- ex3_pragm_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

##### Prior only model

```{r model prior only ex3 , include = FALSE}
predictions.prior.only.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

```{r predictions prior only model ex3, include = FALSE}

# ex3_prior_only_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_prior_only_model.rds")

ex3_prior_only_model <- readRDS("../../stats/saves/ex3_prior_only_model.rds")

# model predictions
ex3_prior_only_model_pred <- ex3_prior_only_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(model="prior_only")

## noise parameter
ex3_prior_only_model_noise <- ex3_prior_only_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

##### Flat prior model

```{r model flat prior ex3, include = FALSE}
predictions.flat.prior.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

  return  [conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment, Math.exp(noisyModelPredictions.score(1))]

```{r predictions prior only model ex3, include = FALSE}

# ex3_flat_prior_model <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model.rds")

ex3_flat_prior_model <- readRDS("../../stats/saves/ex3_flat_prior_model.rds")

# model predictions
ex3_flat_prior_model_pred <- ex3_flat_prior_model %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment"), sep="/")%>%
  mutate(model="flat_prior")

## noise parameter
ex3_flat_prior_model_noise <- ex3_flat_prior_model %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

# summarize data 
adult_ex3_data_summary <- adult_ex3_data %>%
  mutate(model = "data") %>%
  group_by(model,common_ground_manipulation,speaker,alignment)%>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         ci_lower  = qbeta(.025, a, b),
         ci_upper = qbeta(.975, a, b),
         mean = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

# summarize model predictions
ex_3_model_pred <- bind_rows(
  ex3_pragm_model_pred,
  ex3_prior_only_model_pred,
  ex3_flat_prior_model_pred
)%>%
  group_by(model,common_ground_manipulation,speaker,alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper = hdi_upper(value))

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  adult_ex3_data_summary
)

##### Plot model predictions and data

```{r plot ex3 model predictions and data}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","flat_prior","prior_only"),
                     values= c("black","#859900","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

##### Correlation plot model predictions and data

```{r plot }

ex_3_cor_plot <- ex_3_pred %>%
  ungroup()%>%
  filter(model != "data") %>%
  mutate(data_mean = rep(ex_3_pred%>%filter(model == "data")%>%pull(mean),3),
         data_ci_lower = rep(ex_3_pred%>%filter(model == "data")%>%pull(ci_lower),3),
         data_ci_upper = rep(ex_3_pred%>%filter(model == "data")%>%pull(ci_upper),3))

ggplot(data = ex_3_cor_plot,aes(x = mean, y = data_mean, col = "#BB5566")) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = 1)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0,size = 1)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  guides(col = F)+
  facet_grid ( ~ model) +
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = mean, y = data_mean), inherit.aes = F, size = 3)+
  theme_few()

#### Model comparison

```{r model likelihood rsa ex3, include = FALSE}
loglike.rsa.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

 return sum(conditionOutput)

```{r model likelihood prior only ex3 , include = FALSE}
loglike.prior.only.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

   var conditionSpecificPriors =  _.filter(allPriorProbs, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

 return sum(conditionOutput)

```{r model likelihood flat prior ex3, include = FALSE}
loglike.flat.prior.ex3 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

  var conditionSpecificData =  _.filter(allData, {common_ground_manipulation: conditionInfo.common_ground_manipulation, speaker: conditionInfo.speaker, alignment: conditionInfo.alignment})

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

 return sum(conditionOutput)

```{r ex3 model comparison}
# ex3_pragm_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex3_pragm_model_loglike.rds")

ex3_pragm_model_loglike <- readRDS("../../stats/saves/ex3_pragm_model_loglike.rds")%>%
  mutate(model = "pragmatic")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex3_prior_only_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.prior.only.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# )%>%
#   saveRDS("../../stats/saves/ex3_prior_only_model_loglike.rds")

ex3_prior_only_model_loglike <- readRDS("../../stats/saves/ex3_prior_only_model_loglike.rds")%>%
  mutate(model = "prior_only")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex3_flat_prior_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.flat.prior.ex3, sep='\n'),
#   data = list(allData = adult_ex3_data, allConditions = allConditionsEx3, predPrior = predPriorEx3),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# )%>%
#   saveRDS("../../stats/saves/ex3_flat_prior_model_loglike.rds")

ex3_flat_prior_model_loglike <- readRDS("../../stats/saves/ex3_flat_prior_model_loglike.rds")%>%
  mutate(model = "flat_prior")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

ex3_model_comparison <- bind_rows(
  ex3_pragm_model_loglike,
  ex3_prior_only_model_loglike,
  ex3_flat_prior_model_loglike
)%>%
  mutate(model = reorder(model,logP))
```
##### Plot for likelihoods

```{r}
ggplot(ex3_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=0, yend=logP), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#859900","#b58900","#6c71c4"))
```

##### Bayes factors

```{r}
ex3_bf <- data.frame(
  comparison = c("pragmatic vs. prior_only","pragmatic vs. flat_prior", "flat_rior vs. prior_only"),
  bayes_factor = c(
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "pragmatic"] - ex3_model_comparison$logP[ex3_model_comparison$model== "prior_only"]),
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "pragmatic"] - ex3_model_comparison$logP[ex3_model_comparison$model== "flat_prior"]),
    exp(ex3_model_comparison$logP[ex3_model_comparison$model== "prior_only"] - ex3_model_comparison$logP[ex3_model_comparison$model== "flat_prior"]))
)

ex3_bf%>%
  kable(digits = 2)
```

@@

```{r model setup ex 4, include = FALSE}

@@

# combining all priors into one data frame
predPriorEx4 = rbind(strongPrior,mediumPrior,weakPrior)

# data frame to keep track of unique conditions
allConditionsEx4 <- predPriorEx4 %>%
  select(-prior)

@@

#### Model predictions

##### Pragmatic model

```{r model rsa noise ex4, include = FALSE}
predictions.rsa.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)

    return  extend([conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment+ "/" + conditionInfo.prior_manipulation, Math.exp(noisyModelPredictions.score(1))])

   conditionOutput

'
```

ex4_pragm_model <- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4, sep='\n'),
  data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
  data_var = "dataFromR"
)

# model predictions
ex4_pragm_model_pred <- ex4_pragm_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(value = `1`,
         model="pragmatic",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))%>%
  select(-`1`)

##### Prior only model

```{r model prior only ex4 , include = FALSE}
predictions.prior.only.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

  return  extend([conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment+ "/" + conditionInfo.prior_manipulation, Math.exp(noisyModelPredictions.score(1))])

conditionOutput

'
```

```{r predictions prior only model ex 4, include = FALSE}

ex4_prior_only_model <- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex4, sep='\n'),
  data = list(allData = adult_ex4_data, allConditions = allConditions, predPrior = predPrior, noise = ex3_prior_only_model_noise$mean),
  data_var = "dataFromR"
)

# model predictions
ex4_prior_only_model_pred <- ex4_prior_only_model %>%
  separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(value = `1`,
         model="prior_only",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))%>%
  select(-`1`)

##### Flat prior model

```{r model flat prior ex4, include = FALSE}
predictions.flat.prior.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

  return  extend([conditionInfo.common_ground_manipulation + "/" + conditionInfo.speaker + "/" +  conditionInfo.alignment+ "/" + conditionInfo.prior_manipulation, Math.exp(noisyModelPredictions.score(1))])

conditionOutput

'
```

```{r predictions flat prior model ex 4, include = FALSE}

ex4_flat_prior_model <- webppl(
  program_code = paste(rsaUtils, rsaModel, predictions.flat.prior.ex4, sep='\n'),
  data = list(allData = adult_ex4_data, allConditions = allConditions, predPrior = predPrior, noise = ex3_flat_prior_model_noise$mean),
  data_var = "dataFromR"
)

# model predictions
ex4_flat_prior_model_pred <- ex4_flat_prior_model %>%
separate(`0`, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(value = `1`,
         model="flat_prior",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))%>%
  select(-`1`)
```

# summarize model predictions
ex_4_model_pred <- bind_rows(
  ex4_pragm_model_pred,
  ex4_prior_only_model_pred,
  ex4_flat_prior_model_pred
) %>%
  group_by(model,common_ground_manipulation,prior_manipulation,speaker,alignment)%>%
  rename(mean = value)%>%
  summarise(mean = mean,
            ci_lower = mean,
            ci_upper= mean)

##### Plot model predictions and data

##### Correlation plot model predictions and data

ggplot(data = ex_4_cor_plot,aes(x = mean, y = data_mean, col = prior_manipulation)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = 1)+
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0,size = 1)+
  coord_fixed()+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  facet_grid ( ~ model) +
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = mean, y = data_mean), inherit.aes = F, size = 3)+
  theme_few(base_size = ) + 
  scale_colour_ptol()
```

#### Model comparison

```{r model likelihood rsa ex4, include = FALSE}
loglike.rsa.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

var model = function(){

    var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].prior)

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

  return sum(conditionOutput)

```{r model likelihood prior only ex4 , include = FALSE}
loglike.prior.only.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

var model = function(){

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

  return sum(conditionOutput)
}
'
```

```{r model likelihood flat prior ex4, include = FALSE}
loglike.flat.prior.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior
var noise = dataFromR.noise[0]

var model = function(){

    var loglike = Binomial({
        p: Math.exp(noisyModelPredictions.score(1)), 
        n: conditionSpecificData.length}).score(
      sum(_.map(conditionSpecificData, "correct_inf"))
    )

  return  loglike

  return sum(conditionOutput)
}
'
```

```{r ex 4 model comparison}
# ex4_pragm_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.rsa.ex4, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_loglike.rds")

ex4_pragm_model_loglike <- readRDS("../../stats/saves/ex4_pragm_model_loglike.rds")%>%
  mutate(model = "pragmatic")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex4_prior_only_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.prior.only.ex4, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_loglike.rds")

ex4_prior_only_model_loglike <- readRDS("../../stats/saves/ex4_prior_only_model_loglike.rds")%>%
  mutate(model = "prior_only")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

# ex4_flat_prior_model_loglike <- webppl(
#   program_code = paste(rsaUtils, rsaModel, loglike.flat.prior.ex4, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4, noise = ex3_pragm_model_noise$mean),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4, 
#   cores = 4,
#   inference_opts = list(method = "forward", samples = 5000)
# ) %>%
#   saveRDS("../../stats/saves/ex4_flat_prior_model_loglike.rds")

ex4_flat_prior_model_loglike <- readRDS("../../stats/saves/ex4_flat_prior_model_loglike.rds")%>%
  mutate(model = "flat_prior")%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))

ex4_model_comparison <- bind_rows(
  ex4_pragm_model_loglike,
  ex4_prior_only_model_loglike,
  ex4_flat_prior_model_loglike
)%>%
  mutate(model = reorder(model,logP))
```

##### Plot for likelihoods

```{r}
ggplot(ex4_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=0, yend=logP), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#859900","#6c71c4"))
```

##### Bayes factors

```{r}
ex4_bf <- data.frame(
  comparison = c("pragmatic vs. prior_only","pragmatic vs. flat_prior", "flat_rior vs. prior_only"),
  bayes_factor = c(
    exp(ex4_model_comparison$logP[ex4_model_comparison$model== "pragmatic"] - ex4_model_comparison$logP[ex4_model_comparison$model== "prior_only"]),
    exp(ex4_model_comparison$logP[ex4_model_comparison$model== "pragmatic"] - ex4_model_comparison$logP[ex4_model_comparison$model== "flat_prior"]),
    exp(ex4_model_comparison$logP[ex4_model_comparison$model== "flat_prior"] - ex4_model_comparison$logP[ex4_model_comparison$model== "prior_only"]))
)

ex4_bf%>%
  kable(digits = 2)
```
#### Appendix: Model predictions with inferred noise

```{r model rsa noise ex4, include = FALSE}
predictions.rsa.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

@@

```{r predictions pragmatic model ex 4 inf noise, include = FALSE}

# ex4_pragm_model_inf_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.rsa.ex4, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditionsEx4, predPrior = predPriorEx4),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_pragm_model_inf_noise.rds")

ex4_pragm_model_inf_noise <- readRDS("../../stats/saves/ex4_pragm_model_inf_noise.rds")

# model predictions
ex4_pragm_model_pred_inf_noise <- ex4_pragm_model_inf_noise %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(model="pragmatic",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))

## noise parameter
ex4_pragm_model_noise_inf_noise <- ex4_pragm_model_inf_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

```{r model prior only ex4 , include = FALSE}
predictions.prior.only.ex4 <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.predPrior

@@

```{r predictions prior only model ex 4 inf noise, include = FALSE}

# ex4_prior_only_model_inf_noise <- webppl(
#   program_code = paste(rsaUtils, rsaModel, predictions.prior.only.ex4, sep='\n'),
#   data = list(allData = adult_ex4_data, allConditions = allConditions, predPrior = predPrior),
#   data_var = "dataFromR",
#   model_var = "model",
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 500, verbose = T)
# ) %>%
#   saveRDS("../../stats/saves/ex4_prior_only_model_inf_noise.rds")

ex4_prior_only_model_inf_noise <- readRDS("../../stats/saves/ex4_prior_only_model_inf_noise.rds")

# model predictions
ex4_prior_only_model_pred_inf_noise <- ex4_prior_only_model_inf_noise %>%
  filter(!(Parameter %in% c("noise")))  %>%
  separate(Parameter, into = c("common_ground_manipulation", "speaker", "alignment", "prior_manipulation"), sep="/")%>%
  mutate(model="prior_only",
         prior_manipulation = relevel(as.factor(prior_manipulation), ref = "weak"))

## noise parameter
ex4_prior_only_model_noise_inf_noise <- ex4_prior_only_model_inf_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

@@

```{r predictions flat prior model ex 4 inf noise, include = FALSE}
